\chapter{Built-in Procedures and Functions}\index{Procedures}\index{Functions}

Procedures and functions may take parameters (often called {\em
arguments}\index{Arguments}), and these must be enclosed in brackets.
Functions return a value while procedures don't.

\begin{description}
\item[{\tt NUMFORMAT (total, decimals)}]\index{NUMFORMAT}
This affects all future number printing via the {\tt PRINT} statement. The
numbers will be printed right-justified in a total fields width of {\tt
total} with {\tt decimals} digits after the decimal point.
\item[{\tt WAIT (time)}]\index{WAIT}
This waits for {\tt time} seconds. This may be a fractional number,
but the accuracy will depend on the computer you are running it on,
however delays down to 1/100th of a second should be achievable.
\item[{\tt STOP}]\index{STOP}
Program execution is stopped with an message indicating the current
line number.
\item[{\tt END}]\index{END}
Program execution is ended.
\item[{\tt CLS}]\index{CLS}
The screen is cleared.
\item[{\tt HVTAB (x, y)}]\index{HVTAB}
The cursor is moved to the supplied column {\tt x} and line {\tt y}. Note
that (0,0) is top-left of the text screen.
\item[{\tt DEG}]\index{DEG}
Switches the internal angle system to degrees. There are 360 degrees in
a full circle.
\item[{\tt RAD}]\index{RAD}
Switches the internal angle system to radians. There are $2\pi$ radians
in a full circle.
\item[{\tt CLOCK}]\index{CLOCK}
Switches the internal angle system to clock mode. There are 60 minutes
in a full circle.
\item[{\tt SWAP (v1, v2)}]\index{SWAP}
This swaps the value of the 2 variables round. Both arguments must be
the same type - ie.  both numberic or both string.
\end{description}
